### 2025-03-20 article_bihom.tex
first commit

% !TXS template
\documentclass[french]{article}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage[a4paper]{geometry}
\usepackage{babel}
\usepackage{amsmath}
\usepackage{amsfonts}
\begin{document}
Ce que j'ai compris dans l'article: Gröbner bases of bihomogeneous ideals generated by polynomials of bidegree (1, 1): Algorithms and complexity de Faugère et al.
\newline

\section{Définitions et notations}
\begin{itemize}
\item $I_{i-1}:f_i$ is spanned by $I_{i-1}$ and by the maximal minors of $\text{jac}_{\textbf{x}}(F_{i-1})$ (if $i > n_y + 1$) and $\text{jac}_{\textbf{y}}(F_{i-1})$ (if $i > n_x + 1$).
\end{itemize}

\end{document}